\large{\bf Large effects and the infinitesimal model }

### 2023-07-07 writeup.tex
.

Stable distributions are by no means new to population biology.
For instance, it is a long-standing question how often the motions of organisms
are better modeled by essentially Brownian random walks or by L\'evy processes \citep{benhamou2007animals}
and a growing literature studies the effects of such power law distributions
on spatial genetic and dynamics
\citep[e.g.,][]{paulose2019spatial,smith2023isolation}.
More directly analogous to the problem we study here, \citet{landis2012phylogenetic}
found that an $\alpha$-stable process (with $\alpha \approx 1.5$) fit the evolution of log endocranial volume
on the primate phylogeny
better than either a Brownian model.
It is also important to note that
the infinitesimal model implies a Gaussian distribution for the difference between an offspring's trait and their parents' average,
\emph{not} for the entire population.
The question of when the trait distribution in an evolving population
might be modeled as a Gaussian has been discussed extensively
\citep[e.g.,][]{turelli1994genetic,dbarre2015evolution}.

@@

\paragraph{Reproductive Terms:}
First consider the limit
\[
		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right).
\]
Because $\psi(\theta)$ is the logarithm of a characteristic function, it is continuous in some neighbourhood of, and vanish at 0.
If $\phi$ is not differentiable at zero (and so the corresponding distribution
does not have a well-defined mean),
the limit does not exist.
On the other hand if $\psi$ is differentiable at zero,
then $\alpha$ is continuous there,
and for all integers $k$, 
\[
	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
\]
In particular, we see that given any continuous function $\alpha(\theta)$, $\theta \in [1,2]$, such that $\alpha(2) = 2\alpha(1)$, $\alpha$ is defined for all positive reals.  We can extend the definition to the negative reals via the Hermitian property for $\phi(\theta)$: $\phi(-\theta) = \overline{\phi(\theta)}$ implies that $\alpha(-\theta) = \overline{\alpha(\theta)}$.

@@

\begin{figure}
    \begin{center}
        \includegraphics{sims/neutral_seg_noise_small}
    \end{center}
    \caption{
        As in Figure~\ref{fig:seg_noise},
        except computed using only mutations with absolute value effect size less than 0.1.
        The same quantities for the simulations of Figure~\ref{fig:sel_seg_noise}
        with stabilizing selection are shown in Figure~\ref{fig:sel_seg_noise_small}.
        \label{fig:seg_noise_small}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics{sims/selected_seg_noise_small}
    \end{center}
    \caption{
        As in Figure~\ref{fig:seg_noise},
        except computed using only mutations with absolute value effect size less than 0.1.
        \label{fig:sel_seg_noise_small}
    }
\end{figure}

>>>>

Stable distributions are by no means new to population biology.
For instance, it is a long-standing question how often the motions of organisms
are better modeled by essentially Brownian random walks or by L\'evy processes \citep{benhamou2007animals}
and a growing literature studies the effects of such power law distributions
on spatial genetic and dynamics \comment{I'm not quite sure what you mean by that}
\citep[e.g.,][]{paulose2019spatial,smith2023isolation}.
More directly analogous to the problem we study here, \citet{landis2012phylogenetic}
found that an $\alpha$-stable process (with $\alpha \approx 1.5$) fit the evolution of log endocranial volume
on the primate phylogeny
better than either a Brownian model.
It is also important to note that
the infinitesimal model implies a Gaussian distribution for the difference between an offspring's trait and their parents' average,
\emph{not} for the entire population.
The question of when the trait distribution in an evolving population
might be modeled as a Gaussian has been discussed extensively
\citep[e.g.,][]{turelli1994genetic,dbarre2015evolution}.

@@

\paragraph{Reproductive Terms:}
First consider the limit
\[
		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right).
\]
Because $\psi(\theta)$ is the logarithm of a characteristic function, it is continuous in some neighbourhood of, and vanishes at 0.
If $\phi$ is not differentiable at zero (and so the corresponding distribution
does not have a well-defined mean),
the limit does not exist.
On the other hand if $\psi$ is differentiable at zero,
then $\alpha$ is continuous there,
and for all integers $k$, 
\[
	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
\]
In particular, we see that given any continuous function $\alpha(\theta)$, $\theta \in [1,2]$, such that $\alpha(2) = 2\alpha(1)$, $\alpha$ is defined for all positive reals.  We can extend the definition to the negative reals via the Hermitian property for $\phi(\theta)$: $\phi(-\theta) = \overline{\phi(\theta)}$ implies that $\alpha(-\theta) = \overline{\alpha(\theta)}$.

@@

\begin{figure}
    \begin{center}
        \includegraphics{sims/neutral_seg_noise_small}
    \end{center}
    \caption{
        As in Figure~\ref{fig:seg_noise},
        except computed using only mutations with absolute value effect size less than 0.2.
        The same quantities for the simulations of Figure~\ref{fig:sel_seg_noise}
        with stabilizing selection are shown in Figure~\ref{fig:sel_seg_noise_small}.
        \label{fig:seg_noise_small}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics{sims/selected_seg_noise_small}
    \end{center}
    \caption{
        As in Figure~\ref{fig:seg_noise},
        except computed using only mutations with absolute value effect size less than 0.2.
        \label{fig:sel_seg_noise_small}
    }
\end{figure}
